R Notebook

### 2021-11-02 FEM_R/Wealth_Health_Socioeconomics.Rmd
Added some code to check the initial conditions of the repl_standard population (eventually will collect into function to run for each initial population). Will also eventually look at these specifics by wealth_quintile. Also used ggplot to generate BMI histograms and overlay

# Handling Stata data files
require(haven)
require(foreign)

@@

rm(repl.final, repl.final2, repl2, absorbing_vars, change.vars, demog.vars, id.var)
```

@@

>>>>

# Handling Stata data files
require(haven)

@@

rm(repl.final, repl.final2, repl2, absorbing_vars, change.vars, demog.vars, id.var, repl)
```

@@

### Initial Conditions

We want to check the initial conditions of each population (or at least for each quintile). Focus on these variables first:
- Gender
- Risk behaviours (Smoking, Drinking, BMI, Exercise status)
- ADLs/IADLs (Disabled?)
- Severe/Mild Condition?

```{r "summary output FAIL"}
cohort <- read_dta('output/WEALTH/ELSA_cohort/ELSA_cohort_summary.dta')

### For first run through of this lets look at the summary output. Half of the work is already done for us here

select.vars <- c('male', 'smoke', 'drink', 'exstat', 'bmi', 'disabled', 'disease', 'Condition')

## To check the initial conditions we only need 2012 info (as this is before simulation)
# Then remove the _all _m & _f columns as we only want outputs by wealth quintile
init.c <- cohort %>% filter(year == 2012) %>%
                      select(-ends_with(c('_all', '_m', '_f'))) %>%
                      select(year, contains(all_of(select.vars))) %>%
                      select(-starts_with(c('i', 'n')))
# remove the p_ prefix
colnames(init.c) <- gsub('p_', '', colnames(init.c))
# now pivot to get data by quintile
init.c.pivot <-  pivot_longer(data = init.c,
                              cols = male_wq1:severeCondition_wq5,
                              names_to = c('Variable', 'Quintile'),
                              names_sep = '_w',
                              values_to = 'Prevalence')

## Summary output is actually making this a bit more difficult than it has to be
# Instead lets look at the repl populations and make a function to do all this

rm(init.c, init.c.pivot, cohort)
```

```{r "repl_populations"}

repl <- read_dta('input_data/ELSA_repl_standard.dta')
repl <- zap_labels(repl)
repl.2012 <- repl %>% filter(year == 2012)

## First step is to group by wealth_quint and then start producing summary stats
repl.sum <- repl.2012 %>%
                group_by(wealth_quint) %>%
                summarise(
                  count = n(),
                  mean_age = mean(age),
                  mean_male = mean(male),
                  smoken = mean(smoken),
                  smokev = mean(smokev),
                  heavy_smoker = mean(heavy_smoker),
                  mean_BMI = mean(bmi, na.rm = TRUE),
                  sd_BMI = sd(bmi, na.rm = TRUE),
                  missing_BMI_percent = ((sum(is.na(bmi)) / n()) * 100)
                )

## histogram of BMI by quintile
repl.bmi <- repl.2012 %>% select(wealth_quint, bmi) %>%
                          filter(wealth_quint %in% c(1,3))
repl.bmi <- drop_na(repl.bmi)
repl.bmi$wealth_quint <- as.factor(repl.bmi$wealth_quint)

ggplot(repl.bmi, aes(x = bmi, color = wealth_quint)) +
  geom_histogram()
```
